---
title: "class10"
author: "Grant Reeves"
format: html
editor: visual
---

## Class 10: Structural Bioinformatics 2 

# Comparative structure analysis of Adenylate Kinase
```{r}
library(ggplot2)
```

```{r}
# Install packages in the R console NOT your Rmd/Quarto file

#install.packages("bio3d")
#install.packages("devtools")
#install.packages("BiocManager")

#BiocManager::install("msa")
#devtools::install_bitbucket("Grantlab/bio3d-view")
```

>Q10. Which of the packages above is found only on BioConductor and not CRAN? 

BiocManager::install("msa") we know this becasue it says BioManager 

>Q11. Which of the above packages is not found on BioConductor or CRAN?: 

devtools::install_bitbucket("Grantlab/bio3d-view") and we know that becasue it has a separate install command 

>Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? 

True

```{r}
library(bio3d)

aa <- get.seq("1ake_A")
aa
```

We can now run BLAST with this sequence 

```{r}
#b <- blast.pdb(aa)

```

```{r}
#hits <- plot(b)
```

```{r}
# List out some 'top hits'

#head(hits$pdb.id)
```

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
# Align and superpose structures

Next we will use the pdbaln() function to align and also optionally fit the identified PDB structures.

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
pdbs
```
we can look at the alignments ourselves just by writing `pdbs`, but thats not really that useful becaue my eyes arent good at interpreting big data like this 

we can plot our sequence allignment like this: 

```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```

Now we have all this data we have downloaded and organized, we should perform some analysis on it, lets do a PCA 

```{r}
# Perform PCA

pc.xray <- pca(pdbs)
plot(pc.xray)
```

This plot shows the results on PCA analysis on Adenylate kinase x-ray structures. Each dot represents one possible PCA.

We can cluster the structures by RMSD 

```{r}
rmsd(pdbs)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

This Plot shows that all the possible PCA analysis results come in to 3 main groups. 


## Further Investigation 

To visualize the major structural variations in the ensemble the function mktrj() can be used to generate a trajectory PDB file by interpolating along a give PC (eigenvector):

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
pc1
```

Now that I have my `pc_1.pdb` file I can open this file, pc_1.pdb, in Mol*. In a in web browser page visit https://molstar.org/viewer/ and “Open Files” from the left control panel selecting .









